Verifying Demonstrations

### 2019-11-15 paper/DBSchnorr-n-bit.tex
Adds \pmod q, fixes unique solutions (2^x-1)

We will start with a special case of the \(l\)-bit problem.
For a session~\(i\), we have two possible responses to the two challenges:
\(s_i = r_i + c\alpha\) and \(s_i' = r_i + c'\alpha\).
Let \(c = 1\) and \(c' = 0\).
We assume that the adversary learns all \(m\) bits of \(s_i\), but only \(l\) 
bits of \(s_i'\).
This special case yields the linear equation system
\begin{align*}
  s_i &= r_i + \alpha
  = s_i'[m-1] 2^{m-1} + \dotsb + s_i'[1] 2^1 + s_i'[0] 2^0 + \alpha \\
  s_i'[j_1] &= b_{j_1} \\
  s_i'[j_2] &= b_{j_2} \\
  \vdots \\
  s_i'[j_l] &= b_{j_l},
\end{align*}
where \(s_i'[j]\) denotes the \(j\)th bit of \(s_i'\), and \(j_1, \dotsc, 
j_l\in \{0, \dotsc, m-1\}\) are indices for the bits chosen by the adversary 
and \(b_{j_1}, \dotsc, b_{j_l}\in \bin\) are the bit values.

@@

\begin{lemma}\label{l-bit-advantage}
  Assume the special case above.
  There are \(m-l\) free variables, each taking a value from \(\bin\).
  This yields \(2^{m-l}\) possible solutions for each session.
\end{lemma}
\begin{proof}
  The result directly follows from the fact that there are \(m-l\) free 
  variables.
  Each free variable can only take two possible values (0 or 1), consequently 
  there are \(2^{m-l}\) possible combinations.
\end{proof}

@@

For any choice of \(c\) and \(c'\), we have
\begin{align}
  \label{eqn:general-c}
  s_i &= r_i + c\alpha \\
  \label{eqn:general-cprime}
  s_i' &= r_i + c'\alpha.
\end{align}
We can transform this system of equations into our special case above by 
Gaussian elimination: add \cref{eqn:general-c} multiplied by \(\frac{c'}{c}\) 
to \cref{eqn:general-cprime}.
This proves the following lemma.

This means that \cref{l-bit-advantage} is also true for the general case.

\paragraph*{The most general \(l\)-bit problem}

We run the protocol \(n\) times and let the adversary learn \(m+l\) bits of the 
result~\(s_i, s_i'\).
This yields the following equation system:
\begin{align*}
  s_0[0]2^0 + \dotsb + s_0[m-1]2^{m-1} &= r_0 + c\alpha \\
  s_0'[0]2^0 + \dotsb + s_0'[m-1]2^{m-1} &= r_0 + c'\alpha \\
  \vdots \\
  s_{n-1}[0]2^0 + \dotsb + s_{n-1}[m-1]2^{m-1} &= r_{n-1} + c\alpha \\
  s_{n-1}'[0]2^0 + \dotsb + s_{n-1}'[m-1]2^{m-1} &= r_{n-1} + c'\alpha,
\end{align*}
where \(m+l\) of \(s_i[j]\)'s and \(s_i'[j']\)'s for each \(i\) are known by 
the adversary.

\begin{theorem}
  If \[l < m-1+\frac{1}{n}-\frac{f(\lambda)}{n},\] then there are at least 
  \(2^{f(\lambda)}\) possible solutions, for some function~\(f\) of the 
  security parameter~\(\lambda\).
\end{theorem}
\begin{proof}
  Consider one instance of the protocol:
  \begin{align*}
    s_i[0]2^0 + \dotsb + s_i[m-1]2^{m-1} &= r_i + c\alpha \\
    s_i'[0]2^0 + \dotsb + s_i'[m-1]2^{m-1} &= r_i + c'\alpha.
  \end{align*}
  This yields \(2m+2\) variables with two equations, thus \(2m\) free variables.
  Let the adversary learn \(m+l\) of the bits, \ie left-hand side \(s_i[j]\)'s or 
  \(s_i'[j']\)'s.
  This leaves \(2m-(m+l) = m-l\) free variables and \(2^{m-l}\) possible 
  solutions.

  Repeat this \(n\) times.
  This yields \(n(2m+1)+1\) variables and \(2n\) equations, thus \(2nm+n+1-2n = 
  2mn - n +1 = n(2m-1)+1\) free variables.
  Let the adversary learn \(m+l\) bits in each round, thus \(n(m+l)\) in total.
  This yields \(n(2m-1)+1 -n(m+l) = 2nm-n+1-nm-nl = nm-n+1-nl = n(m-1-l)+1\) 
  free variables and \(2^{n(m-1-l)+1}\) possible solutions.
  Consequently, if we want \(f(\lambda)\) possible solutions, we have that
  \begin{equation*}
    nm-n-nl+1 > f(\lambda) \iff m-1+\frac{1}{n}-\frac{f(\lambda)}{n} > l.
  \end{equation*}
  Which concludes the proof.
\end{proof}

>>>>

We will start with a special case of the \(l\)-bit problem.
For a session~\(i\), we have two possible responses to the two challenges:
\(s_i = r_i + c\alpha\) and \(s_i' = r_i + c'\alpha\).
Let \(c = 1\) and \(c' = 0\).
We assume that the adversary learns all \(m\) bits of \(s_i\), but only \(l\) 
bits of \(s_i'\).
This special case yields the linear equation system
\begin{align*}
  s_i &= r_i + \alpha
  = s_i'[m-1] 2^{m-1} + \dotsb + s_i'[1] 2^1 + s_i'[0] 2^0 + \alpha &\pmod q\\
  s_i'[j_1] &= b_{j_1} &\pmod 2\\
  s_i'[j_2] &= b_{j_2} &\pmod 2\\
            &\vdots \\
  s_i'[j_l] &= b_{j_l}, &\pmod 2
\end{align*}
where \(s_i'[j]\) denotes the \(j\)th bit of \(s_i'\), and \(j_1, \dotsc, 
j_l\in \{0, \dotsc, m-1\}\) are indices for the bits chosen by the adversary 
and \(b_{j_1}, \dotsc, b_{j_l}\in \bin\) are the bit values.

@@

\begin{lemma}\label{simple-l-bit-advantage}
  Assume the special case above.
  There are \(m-l\) free variables, each taking a value from \(\bin\).
  This yields \(2^{m-l}-1\) unique solutions for each session.
\end{lemma}
\begin{proof}
  The result directly follows from the fact that there are \(m-l\) free 
  variables.
  Each free variable can only take two possible values (0 or 1), consequently 
  there are \(2^{m-l}-1\) unique solutions.
\end{proof}

\paragraph*{Arbitrarily chosen \(c\neq c'\)}

For any choice of \(c\) and \(c'\) such that \(c\neq c'\), we have
\begin{align}
  \label{eqn:general-c}
  s_i &= r_i + c\alpha &\pmod q\\
  \label{eqn:general-cprime}
  s_i' &= r_i + c'\alpha. &\pmod q
\end{align}
We can transform this system of equations into our special case above by 
Gaussian elimination: add \cref{eqn:general-c} multiplied by \(\frac{c'}{c}\) 
to \cref{eqn:general-cprime}.
The multiplicative inverse \(\frac{1}{c}\) exists since \(q\) is prime.
This proves the following lemma.

This means that \cref{simple-l-bit-advantage} is also true for the general case.

@@

We run the protocol \(n\) times and let the adversary learn \(m+l\) bits of the 
result~\(s_i, s_i'\).
This yields the following equation system:
\begin{align*}
  s_0[0]2^0 + \dotsb + s_0[m-1]2^{m-1} &= r_0 + c\alpha &\pmod q\\
  s_0'[0]2^0 + \dotsb + s_0'[m-1]2^{m-1} &= r_0 + c'\alpha &\pmod q\\
  \vdots \\
  s_{n-1}[0]2^0 + \dotsb + s_{n-1}[m-1]2^{m-1} &= r_{n-1} + c\alpha &\pmod q\\
  s_{n-1}'[0]2^0 + \dotsb + s_{n-1}'[m-1]2^{m-1} &= r_{n-1} + c'\alpha, &\pmod 
  q
\end{align*}
where \(m+l\) of \(s_i[j]\)'s and \(s_i'[j']\)'s for each \(i\) are known by 
the adversary.

\begin{lemma}\label{l-bit-advantage}
  If \[l < m-1+\frac{1}{n}-\frac{f(\lambda)}{n},\] then there are at least 
  \(2^{f(\lambda)}-1\) unique solutions, for some function~\(f\) of the 
  security parameter~\(\lambda\).
\end{lemma}
\begin{proof}
  Consider one instance of the protocol:
  \begin{align*}
    s_i[0]2^0 + \dotsb + s_i[m-1]2^{m-1} &= r_i + c\alpha \\
    s_i'[0]2^0 + \dotsb + s_i'[m-1]2^{m-1} &= r_i + c'\alpha.
  \end{align*}
  This yields \(2m+2\) variables with two equations, thus \(2m\) free variables.
  Let the adversary learn \(m+l\) of the bits, \ie left-hand side \(s_i[j]\)'s or 
  \(s_i'[j']\)'s.
  This leaves \(2m-(m+l) = m-l\) free variables and \(2^{m-l}-1\) possible 
  solutions.

  Repeat this \(n\) times.
  This yields \(n(2m+1)+1\) variables and \(2n\) equations, thus \(2nm+n+1-2n = 
  2mn - n +1 = n(2m-1)+1\) free variables.
  Let the adversary learn \(m+l\) bits in each round, thus \(n(m+l)\) in total.
  This yields \(n(2m-1)+1 -n(m+l) = 2nm-n+1-nm-nl = nm-n+1-nl = n(m-1-l)+1\) 
  free variables and \(2^{n(m-1-l)+1}-1\) unique solutions.
  Consequently, if we want at least \(2^{f(\lambda)}-1\) unique solutions, we 
  have that
  \begin{equation*}
    nm-n-nl+1 > f(\lambda) \iff m-1+\frac{1}{n}-\frac{f(\lambda)}{n} > l.
  \end{equation*}
  Which concludes the proof.
\end{proof}
